---
title: "Power_Estimation"
author: "Adam Lang"
date: "7/27/2021"
output:
  html_document:
    toc: true
    toc_float: TRUE
    toc_depth: 4
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(simstudy)
library(ADNIMERGE)
library(dplyr)
library(furniture)
library(knitr)
```

```{r, echo=FALSE, include=FALSE}
source("/Users/adamgabriellang/Desktop/clinical_trial_sim/adsim_funcs_data.R")
```

# Pre-Processing/Merging Data

<br>

Combined ADAS11 scores from ADNI1 and ADNI2/GO/3 tables. Used VISCODE for the ADNI1 data and VISCODE2 for the ADNI2/GO/3 data.
ADAS11 scores in ADNI1 come from TOTAL11 and ADAS11 scores in ADNI2/GO/3 come from TOTSCORE (referenced the ADNI Data Dictionary)

```{r, echo=TRUE, eval=TRUE}
adas_scores_1  <- read.csv("/Users/adamgabriellang/Desktop/clinical_trial_sim/Data/ADASSCORES.csv")
adas_scores_23 <- read.csv("/Users/adamgabriellang/Desktop/clinical_trial_sim/Data/ADAS_ADNIGO23.csv")
```

<br>

ADNI1 VISCODE levels by count
```{r}
table(factor(adas_scores_1$VISCODE))
```

<br> 

ADNI2/GO/3 VISCODE2 levels by count
```{r}
table(factor(adas_scores_23$VISCODE2))
```

<br>

Combining all ADAS Scores
```{r}
adas_scores_1  <- adas_scores_1[,c("RID", "VISCODE",  "TOTAL11")]
adas_scores_23 <- adas_scores_23[,c("RID", "VISCODE2", "TOTSCORE")]
colnames(adas_scores_23) <- c("RID", "VISCODE", "TOTAL11")
adas_scores              <- rbind(adas_scores_1, adas_scores_23)
adas_scores              <- adas_scores[order(adas_scores$RID, 
                                              adas_scores$VISCODE, 
                                              decreasing = FALSE), ]
#dropping blank values and unschedules
drops1           <- which(adas_scores$VISCODE == "")
drops2           <- which(adas_scores$VISCODE == "uns1")

adas_scores      <- adas_scores[-unique(union(drops1, drops2)), ]
```

<br>

VISCODE by count
```{r}
table(factor(adas_scores$VISCODE))
```

<br>

Number of subjects in ADAS data
```{r}
nlevels(factor(adas_scores$RID))
```

<br>

Took demographic data from ADNIMERGE
```{r}
column.keeps             <- c("RID", "DX", "COLPROT", "ORIGPROT", 
                               "VISCODE", "EXAMDATE", "AGE", 
                               "PTGENDER", "PTEDUCAT", 
                               "APOE4", "CDRSB", "MMSE", 
                               "M")
adas_demog       <- ADNIMERGE::adnimerge[,column.keeps]
```

<br>

adnimerge VISCODE by count
```{r}
table(factor(adas_demog$VISCODE))
```

<br>


Number of subjects in both ADAS and ADNIMERGE
```{r}
# of subjects in ADAS data alone
nlevels(factor(adas_scores$RID))
```
```{r}
# of subjects in both
length(intersect(levels(factor(adas_scores$RID)), levels(factor(adas_demog$RID))))
```

<br>

Merged demographics and ADAS data
```{r}
adas_scores$RID     <- factor(adas_scores$RID)
adas_scores$VISCODE <- factor(adas_scores$VISCODE)
adas_demog$RID      <- factor(adas_demog$RID)
adas_demog$VISCODE  <- factor(adas_demog$VISCODE)
adas_merge_demog    <- merge(adas_demog, adas_scores, by=c("RID", "VISCODE"))
```

<br>

Reading AV45 data
```{r}
av45             <- read.csv("/Users/adamgabriellang/Desktop/clinical_trial_sim/Data/UCBERKELEYAV45_01_14_21.csv")
av45.keeps       <- c("RID", "EXAMDATE", "VISCODE2", "SUMMARYSUVR_COMPOSITE_REFNORM")
av45             <- av45[,av45.keeps]
```

<br>



Merging AV45 with ADAS/Demographic data based off closest EXAMDATE by subject. Created a Amyloid Pos/Neg column and a column for validity of Amyloid measurement
```{r, cache=TRUE}
#changing column names to match
colnames(av45)   <- c("RID", "EXAMDATE", "VISCODE", "SUMMARYSUVR_COMPOSITE_REFNORM")
av45$AmyPos      <- rep(NA, nrow(av45))

#creating pos/neg column based on defined threshold for summary suvr
pos              <- which(av45$SUMMARYSUVR_COMPOSITE_REFNORM >= .78)
neg              <- which(av45$SUMMARYSUVR_COMPOSITE_REFNORM < .78)
av45["AmyPos"][pos,]  <- 1
av45["AmyPos"][neg,]  <- 0
adas_merge_demog$RID <- factor(adas_merge_demog$RID)
adas_merge_demog$EXAMDATE <- as.POSIXct(adas_merge_demog$EXAMDATE)
av45$EXAMDATE <- as.POSIXct(av45$EXAMDATE)
av45$RID             <- factor(av45$RID)

#merging with AV45 data with ADAS/Demographic data by closest time
adas_merge_demog_av45 <- MergeSubjectTime(adas_merge_demog, av45, "RID", "EXAMDATE", "EXAMDATE")
adas_merge_demog_av45$timediff <- abs(difftime(adas_merge_demog_av45$EXAMDATE.x, adas_merge_demog_av45$EXAMDATE.y, units = "days"))
adas_merge_demog_av45$timediff <- adas_merge_demog_av45$timediff / 365
adas_merge_demog_av45$valid <- rep(NA, nrow(adas_merge_demog_av45))

#determining validity of AV45 measurement
# <= 6mo time difference is valid
# > 6mo is not 

adas_merge_demog_av45["valid"][which(adas_merge_demog_av45$timediff <= .5), ] <- 1
adas_merge_demog_av45["valid"][which(adas_merge_demog_av45$timediff > .5), ] <- 0
adas_merge_demog_av45 <- adas_merge_demog_av45[order(adas_merge_demog_av45$RID,
                                                     adas_merge_demog_av45$EXAMDATE.x,
                                                     decreasing = FALSE), ]
adas_merge_demog_av45 <- na.omit(adas_merge_demog_av45)
adas_merge_demog_av45$RID <- factor(adas_merge_demog_av45$RID)
```

<br>

Created a months since baseline observation column
```{r}
adas_merge_demog_av45$M <- as.numeric(adas_merge_demog_av45$M)
adas_merge_demog_av45   <- TimeSinceBaseline(adas_merge_demog_av45, timecol = "M")
```

Number of subjects

```{r}
nlevels(factor(adas_merge_demog_av45$RID))
```

Months since baseline by count
```{r, out.width}
table(factor(adas_merge_demog_av45$new_time))
```

<br>

Filtering by baseline observations with valid amyloid measurements

```{r}
rows.t1 <- which(adas_merge_demog_av45$new_time==0 & adas_merge_demog_av45$valid==1)
rows.t2 <- which(adas_merge_demog_av45$new_time==6)
rows.t3 <- which(adas_merge_demog_av45$new_time==12)
rows.t4 <- which(adas_merge_demog_av45$new_time==18)
rows.t5       <- which(adas_merge_demog_av45$new_time==24)
adas_keep     <- adas_merge_demog_av45[c(rows.t1, rows.t2, rows.t3, rows.t4, rows.t5), ]
adas_keep$RID <- factor(adas_keep$RID)
adas_keep     <-  adas_keep[order(adas_keep$RID,
                                          adas_keep$new_time,
                                          decreasing = FALSE), ]

```

<br>

Number of subjects
```{r}
nlevels(factor(adas_keep$RID))
```

<br>

Time by count
```{r}
table(factor(adas_keep$new_time))
```

<br>


# Descriptive Statistics (Baseline)



Only baseline observations. All baseline observations had an amyloid measurement with 6mo of ADAS11 test
```{r, echo=FALSE}
#Baseline cohort

adas_baseline          <- subset(adas_keep, new_time == 0)
adas_baseline$DX       <- factor(adas_baseline$DX)
adas_baseline$APOE4    <- factor(adas_baseline$APOE4)
adas_baseline$PTGENDER <- factor(adas_baseline$PTGENDER)
adas_baseline$COLPROT  <- factor(adas_baseline$COLPROT)
adas_baseline$AmyPos  <- factor(adas_baseline$AmyPos)
adas_baseline$new_time  <- factor(adas_baseline$new_time)
adas_desc <- adas_baseline[,c("DX", "APOE4", "PTGENDER", "AGE", "TOTAL11", 
                              "MMSE", "CDRSB", "PTEDUCAT", "SUMMARYSUVR_COMPOSITE_REFNORM",
                              "AmyPos")]

desc.table <- table1(adas_desc, splitby = ~ DX)$Table1
desc.table <- desc.table[c(1, 6:nrow(desc.table)), ]
kable(desc.table)
```











